### 2016-10-29 nips2016/review/rebuttal_michael.txt
rebuttal michael (2016-07-21)

Todo
* compare with GNN
* read recent spatial approaches
* constructed graph on MNIST
* run on RCV1
* review "?" in text

Acceptance
* oral: ~2%
* spotlight: ~5%
* poster: ~22%

Private

It appears to us that we got good reviews from reviewers who consider themselves
experts or are confident. Less confident reviewers seem to have not understood
the paper, thus its potential. E.g. they ask questions whose answers are in the
paper already.

General

We thank all the reviewers for reading the paper and providing valuable
feedback. The specific comments of each reviewer are addressed below.

We already cleaned up and edited the paper in the hope to make it more
accessible. We'll take into account the feedbacks for the final revision.

We plan to publicly release the code once the paper is published so that people
can reproduce our results, compare with our model or apply it to their problems.

Note that very recent works [post-submission refs?] (posted on arXiv after the
submission deadline) shows the growing interest for a generalization of CNNs to
arbitrary graphs. While both these approaches are spatial, we look forward to
compare with them.

Reviewer 1: comparison with GNN.

GNN does embed the graph into an high-dimensional space.
Limited to combinatorial Laplacian, no normalized?.
Classify graphs (graph-focused) or nodes (node-focused).
Thanks for pointing this up. We'll add this discussion to the final version of
the paper.

First, the goal of GNNs is to classify graphs or nodes, while the proposed
method classifies signals on graphs. Then, each approach could be twisted and we
could treat a graph signal as labels, but then there'll be no pooling.
Our method may be used to classify nodes, by treating a graph signal as a set of
labels, but we don't see how GNNs could be used to classify graph signals.

Note suited for multiple signals on graph. But for features who leave on nodes.

GNNs is designed as a spatial method.

Second, GNNs need to perform as many backpropagation operations as there are
edges and nodes in the graph per learning iteration. Complexity of

Graph neural networks (GNNs) (Scarselli et al., 2009) are a recurrent neural
network architecture defined on graphs. GNNs apply recurrent neural networks for
walks on the graph structure, propagating node representations until a fixed
point is reached. The resulting node representations are then used as features
in classification and regression problems. GNNs support only discrete labels and
perform as many backpropagation operations as there are edges and nodes in the
graph per learning iteration. Gated Graph Sequence Neural Networks modify GNNs
to use gated recurrent units and to output sequences (Li et al., 2015).

One back-propagation per output: each node embedding is an output.

Reviewer 1: rotational invariance.

Your analysis is correct, the proposed filters are isotropic, i.e. they are
oblivious to the orientation of an edge. The learned filters therefore provide
rotational invariance. Whether this is a limitation or an asset depends on the
problem and should be verified empirically, as for any invariance (translation,
scale). In an experiment (not in the paper) where we did randomly rotate (from
-90° to 90°) each MNIST digit (no data augmentation), we obtained 91.47%
accuracy while classical CNNs got 70.65%. Rotational invariance can therefore be
an asset. Moreover, data augmentation schemes have used rotated versions of
(part of) images to force a network to learn this invariance, and Spatial
Transformer Networks [1], who learn invariance to translation, scale, rotation
and more generic warping, did show state-of-the-art performance on several
benchmarks. We did not include these preliminary results in the paper as we want
to test on larger image datasets the impact of this invariance. As part of this
future work, we'll investigate the possibility to build our filters on
anisotropic Laplacians [2], which should allow us to learn directional /
oriented filters.

Reviewer 3.

The limited results in Table 4 could be due to the small size (~19k documents)
of the 20NEWS dataset. We performed the same task on RCV1 (800k documents), and
obtained...
Unfortunately after the submission deadline. Or say that Bruna did it.

Reviewer 6.

Thanks for your encouragement. Please note that most of the recent works are
based on a spatial formulation. What are the limitations? We look forward to
compare all these approaches.

Reviewer 7: single graph.

As you rightfully pointed out, our spectral construction does not transfer
directly to other graphs. That is an advantage of spatial constructions?. We do
however believe that graphs from a certain class (e.g. social graphs) share a
similar spectrum, which would allow to learn and transfer the Chebyshev
coefficients across graphs. That is a planned future work.?

Reviewer 7: pooling on graphs.

A pooling strategy on arbitrary graphs will always? require some sort of
coarsening, i.e. a way to group similar vertices. As we pointed out, there
exists plenty of methods for multi-level clustering. The proposed CNN on graphs
is agnostic to the choice of a particular algorithm, and we hope that future
research in the area will bring better algorithms. Moreover, the coarsening step
only has to be performed once, before training.

Reviewer 7: MNIST experiment.

You are right, we did experiment with a regular grid on MNIST because it is the
sole way to compare with classical CNNs. The ~3% performance loss you mention
(from Table 3 I guess?) is caused by the use of random graph, i.e. random
connections between pixels. It is clear that in this setting we expect to loose
much performance, which indeed shows that the grid structure is important for
images. We could also build a graph from the pixel values only, which would
approach a regular grid?. Performances are obviously not state-of-the-art, we
only wanted to compare with the classical LeNet-5 and not bother with model
tuning.

Reviewer 7 & 8: 20NEWS experiment, usefulness of the graph.

We did compare with baselines only. The small performance improvement is
probably because the constructed graph is not good enough. There is probably a
better way. Constructing graphs from data is an open research question. This is
however a non-issue when the data naturally lies on a graph structure. Then why
didn't we do it ??

The usefulness of the graph structure obviously depends on the quality of the
graph. That is the point we tried to make in section 6.2: if a given graph does
not improve much on a random graph, then this graph is not well suited for the
problem at hand. If the graph is well constructed, e.g. the grid for MNIST, then
results are good. Moreover, the very recent works in this area [refs] suggest
that this is a research direction of interest.

Reviewer 8 & 9: understanding.

As you pointed out, you do not seem to be familiar with the field of Graph
Signal Processing. Due to space constraints, we had to write densely and
specifically pointed out a review of the field [29]. Moreover, most of the
reviewers seem to have understood the paper well. We did however clean the paper
up and hope it is easier to read.

Reviewer 8: performance.

Our method is both faster and more effective, see Table 6 (Spline filters), than
[13]. The theoretical difference is that our filters are exactly localized in K
hops.

Reviewer 9: natural graphs.

The problem with signals on natural graphs is that, at least for now, not many
datasets are public, and when they are, they are quite small. ?? We do however
plan to make experiments on Wikipedia, using hyperlinks and user activity to
build a graph. This will however require us to build a new dataset.

[1] Spatial Transformer Networks 
[2] anisotropic 
[3] Rotation-invariant convolutional neural networks for galaxy morphology prediction